```{r librerias}
library(easyPubMed)
library(pubmed.mineR)
```

Descargando ejemplo filtrado de articulos de endometriosis posteriores a 2019:
```{r batch download, eval=FALSE, echo=TRUE}
query <- "endometriosis AND 2020/01/01:3000/12/31[dp]"

batch_pubmed_download(pubmed_query_string = query,
                      dest_dir = "intermediateData/",
                      dest_file_prefix = "last_endometriosis",
                      format = "abstract",
                      batch_size = 5000
                      )

```

Generando objeto clase 'Abstracts':
```{r ejemplo readbs, eval=FALSE, echo=TRUE}
last_abstracts <- readabs("intermediateData/last_endometriosis01.txt")

save(last_abstracts,
     file = "intermediateData/last_abstracts.RData")
```

Ejemplo extraccion codigos de genes:
```{r ejemplo gene atomization, eval=FALSE, echo=TRUE}
last_genes <- gene_atomization(last_abstracts)

save(last_genes,
     file="intermediateData/last_genes.RData")
```

Ejemplo extraccion de palabras:
```{r word atomizations example, eval=FALSE, echo=TRUE}
last_words <- word_atomizations(last_abstracts)

save(last_words,
     file = "intermediateData/last_words.RData")
```

Ejemplo TDM:
```{r term document matrix example, eval=FALSE, echo=TRUE}
tdm <- tdm_for_lsa(last_abstracts, 
                   c("age", "gender", "woman", "women", " man ", 
                     " men ", "smoking", "relationships", "relation"))

save(tdm,
     file = "intermediateData/last_tdm.RData")
```




Descargando todos los artículos de endometriosis:
```{r download all endometriosis, eval=FALSE}
batch_pubmed_download("endometriosis",
                      dest_dir = "data/",
                      dest_file_prefix = "total_endometriosis",
                      format = "abstract",
                      batch_size = 5000
                      )
```

Una vez tenemos los archivos, concatenamos todos los datos en uno único:
```{r concatenate text files, eval=FALSE}
# List of files to be added together
files_list <- list.files(path = "data/",
                         pattern = "total",
                         full.names = TRUE) # include path
# Create new file
out_file <- file(description = "intermediateData/todos.txt",
                 open = "w")
# Read each downloaded file and write into final file
for (i in files_list){
  x <- readLines(i)
  writeLines(x, out_file)
}

close(out_file)
```

Genera objeto S4 (corpus primario):
```{r generate abstract object, eval=FALSE}
# Generate the object
abstracts <- readabs("intermediateData/todos.txt")
# Save object
save(abstracts, file = "intermediateData/abstracts.RData")
str(abstracts)
```


Word atomization:
```{r word atomization, eval=FALSE}
words <- word_atomizations(abstracts)
save(words, file = "intermediateData/words.RData")
# Move text file to results directory
file.rename(from = "word_table.txt",
            "results/words.txt")
```

Gene extraction:
```{r gene extraction, eval=FALSE}
genes <- gene_atomization(abstracts)
save(genes, file = "intermediateData/genes.RData")
# Move text file to results directory
file.rename(from = "table.txt",
            "results/genes.txt")
```

En qué frases estás los genes sospechosos (AMH en concreto):
```{r frases con genes, official_fn, eval=FALSE}
# Returns a text file with the sentences that contain official gene symbols from HGNC.
official_fn(genes = genes,
            abs = abstracts,
            filename = "intermediateData/sus.txt",
            terms = "*")

file.rename(from = "intermediateData/sus.txtofficial.txt",
            "results/sus.txtofficial.txt")

# Structure of the output is:
# >> gene_symbol
# PMID
# Sentence
```

```{r gene extraction pubtator, eval=FALSE}
load("intermediateData/last_abstracts.RData")

# Vector with pmids
pmid_vector <- last_abstracts@PMID[1:5]

# Preallocate list
extraction_list <- vector(mode = "list",
                          length = length(pmid_vector))

# Retrieve genes vectors from PubTator and store in list
for (i in 1:length(pmid_vector)) {
  extraction_list[[i]] <- pubtator_function(pmid_vector[i])$Genes
}

# Add pmid to Gene info in each non-null element of the gene list
for (i in 1:length(extraction_list)) {
if (is.null(extraction_list[[i]]) == FALSE) {
  extraction_list[[i]] <- sub(pattern = "^",
      replacement = paste0(pmid_vector[i],">"),
      extraction_list[[i]])
}
}


# List into vector of genes
genes_vector <- unlist(extraction_list)

# Generate dataframe with pmid, symbol, and geneID in different columns  ###INCOMPLETE ###
genes_dataframe <- data.frame(Symbol = genes_vector,
                              stringsAsFactors = FALSE)
genes_dataframe$GeneID <- sub(pattern = "[[:print:]]*>",
                              replacement = "",
                              genes_dataframe$Symbol)
```


Filtro distribución hipergeométrica
```{r download gene2pubtatorcentral, eval=FALSE}
download.file(url = "ftp://ftp.ncbi.nlm.nih.gov/pub/lu/PubTatorCentral/disease2pubtatorcentral.sample",
              destfile = "data/disease2pubtatorcentral.sample")

```

```{r read gene2pubtatorcentral, eval=FALSE}
gene2pubtator <- read.table(
  file = "../../PEC2_Fase1/PubTatorCentral/gene2pubtatorcentral",
  header = FALSE,
  colClasses = c("integer", "NULL", "character", "NULL", "NULL"),
  col.names = c("PMID", "Type", "GeneID", "Name", "Resource"),
  fill = TRUE,
  quote = "",
  sep = "\t"
)

save(gene2pubtator,
     file = "intermediateData/gene2pubtator.RData")
```

```{r check na values, eval=FALSE}
c(pmid = sum(is.na(gene2pubtator$PMID)),
  geneid = sum(is.na(gene2pubtator$GeneID)))
```

```{r check empty strings, eval = FALSE}
sum(gene2pubtator$GeneID == "")
```

```{r prueba usando vector de indices y muestra, eval=FALSE}
# Load sample
sample_df <- read.table(
  file = "../../PubTator/gene2pubtatorcentral.sample",
  header = FALSE,
  colClasses = c("integer", "NULL", "character", "NULL", "NULL"),
  col.names = c("PMID", "Type", "GeneID", "Name", "Resource"),
  fill = TRUE,
  quote = "",
  sep = "\t"
)

save(sample_df,
     file = "../../PubTator/sample_df.RData")

load("../../PubTator/sample_df.RData")

system.time({
while (TRUE %in% grepl(pattern = ";", sample_df$GeneID)) {
# Retrieve indices
double_geneids_vector <- grep(pattern = ";",
                              x = sample_df$GeneID) 

# Subset dataframe
sample_df_dup <- sample_df[double_geneids_vector, ]

# Borrar primer geneID en df original
sample_df$GeneID <- sub(pattern = "^[0-9]+;",
             replacement = "",
             sample_df$GeneID)

# Conserver primer gene ID en df nuevo
sample_df_dup$GeneID <- sub(pattern = ";[[:graph:]]+",
                             replacement = "", 
                              x = sample_df_dup$GeneID)

# Copiar nueva df a df original
sample_df <- rbind(sample_df,
                   sample_df_dup)
# Borra objetos
rm(list = c("sample_df_dup", 
            "double_geneids_vector"))
}
  sample_df <- unique(sample_df)
  })
```

```{r process double geneID, eval=FALSE}

# Example dataframe to test the method
# dummy_dataframe <- data.frame(PMID = c(123, 456, 789, 789),
#                               GeneID = c("123", "123;234", "567;8910", "8910"),
#                               stringsAsFactors = FALSE
# )

load("intermediateData/gene2pubtator.RData")

round <- 0

while (TRUE %in% grepl(pattern = ";", gene2pubtator$GeneID)) {
# Retrieve indices
double_geneids_vector <- grep(pattern = ";",
                              x = gene2pubtator$GeneID) 

# Subset dataframe
gene2pubtator_dup <- gene2pubtator[double_geneids_vector, ]

# Borrar primer geneID en df original
gene2pubtator$GeneID <- sub(pattern = "^[0-9]+;",
             replacement = "",
             gene2pubtator$GeneID)

# Conserver primer gene ID en df nuevo
gene2pubtator_dup$GeneID <- sub(pattern = ";[[:graph:]]+",
                             replacement = "", 
                              x = gene2pubtator_dup$GeneID)

# Copiar nueva df a df original
gene2pubtator <- rbind(gene2pubtator,
                   gene2pubtator_dup)
# Borra objetos
rm(list = c("gene2pubtator_dup", 
            "double_geneids_vector"))

# Da señales de vida
round <- round + 1
print(paste0("Round ", round,"!" ))
}
# Elimina filas duplicadas
  gene2pubtator <- unique(gene2pubtator)
  })
```

```{r save the deduplicated table, eval = FALSE}
save(gene2pubtator,
     file = "intermediateData/gene2pubtator_dedup.RData")
```

```{r contingency table, eval= FALSE}
load("intermediateData/gene2pubtator_dedup.RData")

# Generate contingency table
# Structure: one-dimensional array
# Each geneID is a column, the name of the column is the geneID
geneID_frequencies <- table(gene2pubtator$GeneID)

save(geneID_frequencies,
     file = "data/geneID_frequencies.RData")
```

```{r check big geneID values, eval = FALSE}
# How did it take 88 rounds to deduplicate geneID field?

load("intermediateData/gene2pubtator.RData")

char <- ";"
string <- gene2pubtator$GeneID

rm("gene2pubtator")

# Leaves only the ; characters
# and counts them for each string in the vector
replaced <- nchar(gsub("[^;]",
                 "",
                 string)
)

# What is the maximum value
maximum <- max(replaced)
# And prints its corresponding string
# may be more than one, but we only retrieve the first maximum
values <- string[which.max(replaced)]

c(ID.quantity = maximum,
  values = values)
```



```{r hypergeometric test, eval=FALSE}
# The hypergeometric test will be used to filter
# the gene results from text mining of the abstracts.
# As in Ji-Long (2016)

# Variables
i <- 721         # Number of retrieved cites containing the geneID
q <- i - 1
m <- 4574        # Total cites containing the geneID
                 # This number is extracted from contingency table
n <- 33e6 - m    # Total cites without the geneID
k <- 29000       # Total number of retrieved cites

# Formula
# Probability for extracting, at least, q results:
P.hyp <- ( 1 - phyper(q, m, n, k))

##
# phyper() gives gives the probability of retrieving q in the sample.
# For a p-value, we want the probability of retrieving AT LEAST q in the sample
# Hence, 1 - phyper()
# I want an over-representation test (probability of more successes than expected).
# H_0: k/n = K/N
# H_A: k/n > K/N

##
# Bonferroni correction
# alpha/nº de tests (en este caso nº de genes)
##

## Implementation of the test
# Variables
gen_df <- as.data.frame(last_genes,
                        stringsAsFactors = FALSE)
gen_df$Freq <- as.integer(gen_df$Freq)
i <- gen_df$Freq[1]
q <- i - 1

```







